Effect of quality-adjusted production function inputs

### 2015-06-03 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Rearranged much of the analysis.

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nests, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
  # Save models to disk.
  save(models, file=fileName)
}
# Put all the coefficients into a data frame
AllCoeffs <- leaf_apply(models, class = c("CDEmodel", "cesModel", "LINEXmodel"), f = function(mod, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  id_list <- strsplit(id, split = "[.]")[[1]]
  Country <- id_list[[1]]
  flavor <- id_list[[2]]
  energy <- id_list[[3]]
  model <- id_list[[4]]
  nest <- if(length(id_list) > 4) id_list[[5]] else NA
  res <- rbind(list(Country=Country, flavor=flavor, energy=energy, model=model, nest=nest))
  coeffs <- naturalCoef(mod)
  res <- cbind(res, coeffs)
  # Now calculate the mean squared error (mse) for each model.
  sqResids <- resid(mod)^2
  mse <- sum(sqResids)/(length(sqResids))
  res <- cbind(res, list(mse=mse))
  return(res)
}) %>% do.call(plyr::rbind.fill, .)
# Get the unadjusted, no-energy values for mse and lambda and place them on each appropriate row
AllCoeffs_unadj_noE <- filter(AllCoeffs, flavor=="unadjusted", energy=="noE") %>% 
                              select(Country, flavor, energy, model, nest, mse, lambda)
colnames(AllCoeffs_unadj_noE)[colnames(AllCoeffs_unadj_noE)=="mse"] <- "mse_unadj_noE"
colnames(AllCoeffs_unadj_noE)[colnames(AllCoeffs_unadj_noE)=="lambda"] <- "lambda_unadj_noE"
AllCoeffs <- merge(AllCoeffs, AllCoeffs_unadj_noE[c("Country", "model", "mse_unadj_noE", "lambda_unadj_noE")])
# Calculate differences relative to the unadjusted, no-energy condition
AllCoeffs <- mutate(AllCoeffs, Dlambda = abs(lambda) - abs(lambda_unadj_noE), Dmse = mse - mse_unadj_noE)

# Relevel factors
AllCoeffs$Country <- relevelFactor(as.factor(unlist(AllCoeffs$Country)), c("PT", "UK"))
AllCoeffs$flavor <- relevelFactor(as.factor(unlist(AllCoeffs$flavor)), c("unadjusted", "adjusted"))
AllCoeffs$energy <- relevelFactor(as.factor(unlist(AllCoeffs$energy)), c("noE", "withE"))
AllCoeffs$model <- relevelFactor(as.factor(unlist(AllCoeffs$model)), c("CD", "CES"))
AllCoeffs$nest <- relevelFactor(as.factor(unlist(AllCoeffs$nest)), c("kl", "kle", "lek", "ekl"))
# Rearrange for easier viewing
AllCoeffs <- arrange(AllCoeffs, Country, model, flavor, energy, nest)
@

@@

This section shows graphs of fits to historical data and residuals.
Because we fit in log-space, the residuals ($r_i$) are defined as
%
\begin{equation} \label{eq:log-residuals}
  r_i \equiv \ln(y_i) - \ln(\hat{y}_i) =\ln\left( \frac{y_i}{\hat{y}_i} \right), 
\end{equation}
%
where $r_i$ will be zero when there is agreement between 
historical~($y_i$) and fitted~($\hat{y}_i$) economic output.

<<Fitted and Residual Graph Functions>>=
fittedGraph <- function(data){
  ggplot() +
    geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=iGDP)) +
    geom_line(data=data, mapping=aes(x=Year, y=fitted, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}

residualPlot <- function(data){
  ggplot() +
    geom_line(data=data, mapping=aes(x=Year, y=Residuals, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    xy_theme()
}
@

<<Graphs CD-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Cobb-Douglas models. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CD"))
@

<<Graphs CD-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for Cobb-Douglas models.">>=
residualPlot(filter(fitted_and_resid_data, model=="CD"))
@

<<Graphs CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($kl$)$e$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="kle" | nest=="kl")))
@

<<Graphs CESkle-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(kl$)$e$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest=="kle" | nest=="kl")))
@

<<Graphs CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($le$)$k$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="lek" | nest=="kl")))
@

<<Graphs CESlek-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with ($le$)$k$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest == "lek" | nest=="kl")))
@

<<Graphs CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($ek$)$l$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="ekl" | nest=="kl")))
@

<<Graphs CESekl-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(ek$)$l$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest=="ekl" | nest=="kl")))
@

%++++++++++++++++++++++++++++++
\subsection{Does quality-adjusting the factors of production 
            decrease the fitting residual and dynamic Solow residual?} 
\label{sec:quality_adj_energy_and_dynamic_solow_residual}
%++++++++++++++++++++++++++++++

We can test our hypothesis that quality-adjusting the factors of production
and including energy will both reduce the Solow residuals~($\lambda$)
and improve the fit to historical data
(thereby, reducing the fitting residuals, $r$)
by calculating 
$\Delta \lambda$ and $\Delta mse$, where
%
\begin{equation} \label{eq:delta_lambda}
  \Delta \lambda \equiv \left| \lambda \right| - \left| \lambda_{Unadjusted, \, Without \, energy} \right|
\end{equation}
%
and
%
\begin{equation} \label{eq:delta_mse}
  \Delta mse \equiv \frac{1}{N} \sum_i{r_i^2} - mse_{Unadjusted, \, Without \, energy}.
\end{equation}
%
where~$N$ is the number of years of data.
When $\Delta \lambda$ and $\Delta mse$ are negative, 
we observe reduction in the Solow residual ($\lambda$) and
the fitting residuals ($r$).

Table~\ref{tab:DSRmse_CD} shows the results for the Cobb-Douglas models.
%
<<MyXTableFunction>>=
myXTable <- function(data, caption, label, digits=2){
  table <- xtable(data, caption=caption, label=label, digits=digits)
  print(table, 
        caption.placement="top", 
        size="\\tiny", 
        include.rownames=FALSE, 
        math.style.negative=TRUE)
}
@
%
<<DSRmse_CD, results="asis">>=
myXTable(filter(AllCoeffs, model=="CD") %>% select(Country, model, flavor, energy, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for Cobb-Douglas models.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
Tables~\ref{tab:DSRmse_CESkle}--\ref{tab:CSRmse_CESekl} show results for the various CES nestings.
%
<<DSRmse_CESkle, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="kle")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($kl$)$e$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
<<DSRmse_CESlek, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="lek")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($le$)$k$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
<<DSRmse_CESekk, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="ekl")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($ek$)$l$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@

%++++++++++++++++++++++++++++++
\subsection{Does quality-adjusting the factors of production 
            decrease the dynamic Solow residual?} 
\label{sec:quality_adj_energy_and_dynamic_solow_residual}
%++++++++++++++++++++++++++++++

@@

Figure~\ref{fig:DynamicSR_CD_Graph} shows the evolution 
of the Solow residual~($\lambda$) over time
as calculated by Equation~\ref{eq:lambdaij} 
with the timespan for finite differences ($t_j - t_i$) set to 
\Sexpr{lag.to.use}~\Sexpr{if (lag.to.use == 1) "year" else "years"}.
It appears that quality-adjusting the factors of production
\emph{and} adding augmenting the production function with energy
decreases the Solow residual~($\lambda$),
i.e., brings the Solow residual closer to zero.

@@

<<DynamicSR_CD_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the Cobb-Douglas model. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CD"))
@

<<DynamicSR_CESkle_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($kl$)$e$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="kle" | nest=="kl")))
@

<<DynamicSR_CESlek_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($le$)$k$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="lek" | nest=="kl")))
@

<<DynamicSR_CESekl_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($ek$)$l$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="ekl" | nest=="kl")))
@

%++++++++++++++++++++++++++++++
\subsection{How does quality-adjusting and including energy affect the dynamic Solow residual?} 
\label{sec:trajectory_graphs}
%++++++++++++++++++++++++++++++

<<SR and R trajectory data>>=
traj.data <- merge(solowResiduals[c("Country", "Year", "model", "flavor", "energy", "nest", "Dlambda_fd")], 
                   fitted_and_resid_data[c("Country", "Year", "model", "flavor", "energy", "nest", "Dr")]) %>%
                   arrange(Country, model, flavor, energy, nest)
traj.data.means <- traj.data %>% 
  group_by(model, Country, flavor, energy, nest) %>%
  summarise(Dlambda_fd.mean = mean(Dlambda_fd, na.rm=TRUE), Dr.mean = mean(Dr))
@

<<SR and R trajectory plot function>>=
residualsTrajGraph <- function(data){
  plot <- ggplot(data=data) + 
    geom_point(mapping=aes(x=Dr, y=Dlambda_fd, color=energy)) + 
#     geom_path(mapping=aes(x=Dr, y=Dlambda_fd, color=energy)) + 
    facet_grid(Country ~ flavor) +
    xlab(expression(Delta*r)) +
    ylab(expression(Delta*lambda)) +
    xy_theme() + 
    theme(legend.title=element_blank())
  return(plot)
}
@

@@

We can also calculate the average values for $\Delta \lambda$ and $\Delta r$ 
as shown in Table~\ref{tab:traj.means.CD}.
%
<<Trajectory_Means_CD, results="asis">>=
myXTable(data = filter(traj.data.means, model=="CD") %>% select(-nest), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for Cobb-Douglas models.",
         label="tab:traj.means.CD",
         digits=6)
@
%
These results for the Cobb-Douglas model indicate that 
we should perform the same analysis for the CES model.
%
<<Solow_residual_trajectories_CESkle, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the CES model with ($kl$)$e$ nesting.">>=
residualsTrajGraph(filter(traj.data, model=="CES" & (nest=="kl" | nest=="kle")))
@
%
<<Trajectory_Means_CESkle, results="asis">>=
myXTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="kle")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($kl$)$e$ nesting.",
         label="tab:traj.means.CESkle",
         digits=6)
@

@@

<<Trajectory_Means_CESlek, results="asis">>=
trajMeansTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="lek")), 
               caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($le$)$k$ nesting.",
               label="tab:traj.means.CESlek",
               digits=6)
@

@@

<<Trajectory_Means_CESekl, results="asis">>=
myXTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="ekl")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($ek$)$l$ nesting.",
         label="tab:traj.means.CESekl",
         digits=6)
@

>>>>

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nests, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
  # Save models to disk.
  save(models, file=fileName)
}
# Put all the coefficients into a data frame
AllCoeffs <- leaf_apply(models, class = c("CDEmodel", "cesModel", "LINEXmodel"), f = function(mod, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  id_list <- strsplit(id, split = "[.]")[[1]]
  Country <- id_list[[1]]
  flavor <- id_list[[2]]
  energy <- id_list[[3]]
  model <- id_list[[4]]
  nest <- if(length(id_list) > 4) id_list[[5]] else NA
  res <- rbind(list(Country = Country, 
                    flavor = if (flavor=="unadjusted") "Unadjusted" else "Quality-adjusted",
                    energy = if (energy=="noE") "Without energy" else "With energy",
                    model=model, 
                    nest=nest))
  coeffs <- naturalCoef(mod)
  res <- cbind(res, coeffs)
  # Now calculate the mean squared error (mse) for each model.
  sqResids <- resid(mod)^2
  mse <- sum(sqResids)/(length(sqResids))
  res <- cbind(res, list(mse=mse))
  return(res)
}) %>% do.call(plyr::rbind.fill, .)
# Get the unadjusted, no-energy values for mse and lambda and place them on each appropriate row
AllCoeffs_unadj_noE <- filter(AllCoeffs, flavor=="Unadjusted", energy=="Without energy") %>% 
                              select(Country, flavor, energy, model, nest, mse, lambda)
colnames(AllCoeffs_unadj_noE)[colnames(AllCoeffs_unadj_noE)=="mse"] <- "mse_unadj_noE"
colnames(AllCoeffs_unadj_noE)[colnames(AllCoeffs_unadj_noE)=="lambda"] <- "lambda_unadj_noE"
AllCoeffs <- merge(AllCoeffs, AllCoeffs_unadj_noE[c("Country", "model", "mse_unadj_noE", "lambda_unadj_noE")])
# Calculate differences relative to the unadjusted, no-energy condition
AllCoeffs <- mutate(AllCoeffs, Dlambda = abs(lambda) - abs(lambda_unadj_noE), Dmse = mse - mse_unadj_noE)

# Relevel factors
AllCoeffs$Country <- relevelFactor(as.factor(unlist(AllCoeffs$Country)), c("PT", "UK"))
AllCoeffs$flavor <- relevelFactor(as.factor(unlist(AllCoeffs$flavor)), c("Unadjusted", "Quality-adjusted"))
AllCoeffs$energy <- relevelFactor(as.factor(unlist(AllCoeffs$energy)), c("Without energy", "With energy"))
AllCoeffs$model <- relevelFactor(as.factor(unlist(AllCoeffs$model)), c("CD", "CES"))
AllCoeffs$nest <- relevelFactor(as.factor(unlist(AllCoeffs$nest)), c("kl", "kle", "lek", "ekl"))
# Rearrange for easier viewing
AllCoeffs <- arrange(AllCoeffs, Country, model, flavor, energy, nest)
@

@@

<<Fitted and Residual Graph Functions>>=
fittedGraph <- function(data){
  ggplot() +
    geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=iGDP)) +
    geom_line(data=data, mapping=aes(x=Year, y=fitted, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
    xlab("") +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}
@

Both historical GDP and fitted GDP are shown in 
Figures~\ref{fig:Graphs CD-fitted}--\ref{fig:Graphs CESekl-fitted}.
%
<<Graphs CD-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Cobb-Douglas models. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CD"))
@
%
<<Graphs CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($kl$)$e$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="kle" | nest=="kl")))
@
%
<<Graphs CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($le$)$k$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="lek" | nest=="kl")))
@
%
<<Graphs CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($ek$)$l$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="ekl" | nest=="kl")))
@

%++++++++++++++++++++++++++++++
\subsection{Fitting residuals} 
\label{sec:fitting_residuals}
%++++++++++++++++++++++++++++++

<<ResidualPlotFunction>>=
residualPlot <- function(data){
  ggplot() +
    geom_line(data=data, mapping=aes(x=Year, y=Residuals, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    xlab("") +
    ylab("r") + 
    xy_theme()
}
@

Because we fit in log-space, fitting residuals ($r_i$) are defined as
%
\begin{equation} \label{eq:log-residuals}
  r_i \equiv \ln(y_i) - \ln(\hat{y}_i) =\ln\left( \frac{y_i}{\hat{y}_i} \right), 
\end{equation}
%
where $r_i$ will be zero when there is agreement between 
historical~($y_i$) and fitted~($\hat{y}_i$) economic output.

The mean squared error~($mse$) for any fitted model can be calculated by
%
\begin{equation} \label{eq:mse}
  mse \equiv \frac{1}{N} \sum_{i=1}^N{r_i^2} .
\end{equation}

Figures~\ref{fig:Graphs CD-resid}--\ref{fig:Graphs CESekl-resid}
show fitting residuals for all models.
%
<<Graphs CD-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for Cobb-Douglas models.">>=
residualPlot(filter(fitted_and_resid_data, model=="CD"))
@
%
<<Graphs CESkle-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with $(kl$)$e$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest=="kle" | nest=="kl")))
@
%
<<Graphs CESlek-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with ($le$)$k$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest == "lek" | nest=="kl")))
@
%
<<Graphs CESekl-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with $(ek$)$l$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest=="ekl" | nest=="kl")))
@

%++++++++++++++++++++++++++++++
\subsection{Dynamic Solow residuals} 
\label{sec:dynamic_solow_residuals}
%++++++++++++++++++++++++++++++

@@

@@

``Dynamic'' Solow residuals~($\lambda_{i,j}$) can be calculated 
for the Cobb-Douglas and CES models
as shown in Appendices~\ref{sec:dynamic_sr_CD} and~\ref{sec:dynamic_sr_CES},
respectively.

Figure~\ref{fig:DynamicSR_CD_Graph} shows the evolution 
of the Solow residual~($\lambda_{i,j}$) over time
as calculated by Equation~\ref{eq:lambdaij} 
with the timespan for finite differences ($t_j - t_i$) set to 
\Sexpr{lag.to.use}~\Sexpr{if (lag.to.use == 1) "year" else "years"}.
%
<<DynamicSR_CD_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the Cobb-Douglas model. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CD"))
@
%
<<DynamicSR_CESkle_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($kl$)$e$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="kle" | nest=="kl")))
@
%
<<DynamicSR_CESlek_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($le$)$k$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="lek" | nest=="kl")))
@
%
<<DynamicSR_CESekl_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($ek$)$l$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="ekl" | nest=="kl")))
@

%++++++++++++++++++++++++++++++
\subsection{Does quality-adjusting the factors of production 
            decrease the mean squared error and/or the Solow residual?} 
\label{sec:quality_adj_energy_and_solow_residual}
%++++++++++++++++++++++++++++++

We can test our hypothesis that quality-adjusting the factors of production
and including energy will both reduce the Solow residuals~($\lambda$)
and improve the fit to historical data
(thereby, reducing the fitting residuals, $r_i$)
by calculating 
$\Delta \lambda$ and $\Delta mse$, where
%
\begin{equation} \label{eq:delta_lambda}
  \Delta \lambda \equiv \left| \lambda \right| - \left| \lambda_{Unadjusted, \, Without \, energy} \right|
\end{equation}
%
and
%
\begin{equation} \label{eq:delta_mse}
  \Delta mse \equiv mse - mse_{Unadjusted, \, Without \, energy}.
\end{equation}
%
where~$N$ is the number of years of data.
When $\Delta \lambda$ and $\Delta mse$ are negative, 
we observe reduction in the Solow residual ($\lambda$) and
the fitting residuals ($mse$)
relative to the unadjusted, no-energy condition.

Table~\ref{tab:DSRmse_CD} shows the results for the Cobb-Douglas models.
%
<<MyXTableFunction>>=
myXTable <- function(data, caption, label, digits=2){
  table <- xtable(data, caption=caption, label=label, digits=digits)
  print(table, 
        caption.placement="top", 
        size="\\tiny", 
        include.rownames=FALSE, 
        math.style.negative=TRUE)
}
@
%
<<DSRmse_CD, results="asis">>=
myXTable(filter(AllCoeffs, model=="CD") %>% select(Country, model, flavor, energy, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for Cobb-Douglas models.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
% 
Tables~\ref{tab:DSRmse_CESkle}--\ref{tab:DSRmse_CESekl} show results 
for CES models with various nestings.
%
<<DSRmse_CESkle, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="kle")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($kl$)$e$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
<<DSRmse_CESlek, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="lek")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($le$)$k$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
<<DSRmse_CESekl, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="ekl")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($ek$)$l$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
Figure~\ref{fig:DSRmse_CD_graph} summarizes the data for the Cobb-Douglas model.
%
<<DSRmse_CD_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Change in Solow Residuals~($\\Delta\\lambda$) and mean squared error~($\\Delta mse$) for the Cobb-Douglas model relative to the unadjusted, no-energy condition.">>=
ggplot(data=filter(AllCoeffs, model=="CD")) +
  geom_point(mapping=aes(x=Dmse, y=Dlambda, color = energy)) +
  facet_grid(Country ~ flavor) +
  xlab(expression(Delta * mse)) + 
  ylab(expression(Delta * lambda)) +
  xy_theme() + 
  theme(legend.title=element_blank())
@
%
Figure~\ref{fig:DSRmse_CES_graph} summarizes the data for the CES models.
%
<<DSRmse_CES_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Change in Solow Residuals~($\\Delta\\lambda$) and mean squared error~($\\Delta mse$) for the Cobb-Douglas model relative to the unadjusted, no-energy condition.">>=
ggplot(data=filter(AllCoeffs, model=="CES")) +
  geom_point(mapping=aes(x=Dmse, y=Dlambda, color = energy, shape=nest)) +
  facet_grid(Country ~ flavor) +
  xlab(expression(Delta * mse)) + 
  ylab(expression(Delta * lambda)) +
  xy_theme() +
  theme(legend.title=element_blank())
@

%++++++++++++++++++++++++++++++
\subsection{Does quality-adjusting the factors of production 
            decrease fitting residuals and/or the dynamic Solow residual?} 
\label{sec:quality_adj_energy_and_dynamic_solow_residual}
%++++++++++++++++++++++++++++++

<<SR and R trajectory data>>=
traj.data <- merge(solowResiduals[c("Country", "Year", "model", "flavor", "energy", "nest", "Dlambda_fd")], 
                   fitted_and_resid_data[c("Country", "Year", "model", "flavor", "energy", "nest", "Dr")]) %>%
                   arrange(Country, model, flavor, energy, nest)
traj.data.stats <- traj.data %>% 
  group_by(model, Country, flavor, energy, nest) %>%
  summarise(Dlambda_fd.mean = mean(Dlambda_fd, na.rm=TRUE), Dr.mean = mean(Dr))

############################################
# Dlambda_fd.p <- traj.data %>% 
#   group_by(model, Country, flavor, energy, nest) %>% 
#   select(Dlambda_fd) %>%
#   t.test(., alternative="less")[["p.value"]]
############################################  
@

<<SR and R trajectory plot function>>=
residualsTrajGraph <- function(data){
  plot <- ggplot(data=data) + 
    geom_point(mapping=aes(x=Dr, y=Dlambda_fd, color=energy)) + 
    facet_grid(Country ~ flavor) +
    xlab(expression(Delta*r[i])) +
    ylab(expression(Delta*lambda[ij])) +
    xy_theme() + 
    theme(legend.title=element_blank())
  return(plot)
}
@

@@

We can also calculate the average values for $\Delta \lambda_{i,j}$ and $\Delta r_{i}$ 
as shown in Table~\ref{tab:Trajectory_Means_CD}.
%
<<Trajectory_Means_CD, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CD") %>% select(-nest), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for Cobb-Douglas models.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@
%
The following tables and figures show the same analysis for the CES model.
%
<<Solow_residual_trajectories_CESkle, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the CES model with ($kl$)$e$ nesting.">>=
residualsTrajGraph(filter(traj.data, model=="CES" & (nest=="kl" | nest=="kle")))
@
%
<<Trajectory_Means_CESkle, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CES" & (nest=="kl" | nest=="kle")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($kl$)$e$ nesting.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@

@@

<<Trajectory_Means_CESlek, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CES" & (nest=="kl" | nest=="lek")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($le$)$k$ nesting.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@

@@

<<Trajectory_Means_CESekl, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CES" & (nest=="kl" | nest=="ekl")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($ek$)$l$ nesting.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@

% %++++++++++++++++++++++++++++++
% \subsection{What effect does including energy and quality-adjustment have 
